---
title: "BAPS Invariance testing Joachim_Kowalski"
author: "Jan Szczypiński"
date: "08 10 2021"
output:
  pdf_document: default
  word_document: default
---

## The MI analysis of BAPS

First loading the packages, data and a function
```{r, echo=FALSE}
library(lavaan)
library(semTools)
source('F:/projekty/robust_diff.R')

setwd("f:/projekty/Joachim/BAPS")

d <- read.csv("BAPS.csv", sep=",", header =T)
```

We create a matrix to compare fit indices across models with different constraints
```{r}

all.results<-matrix(NA, nrow = 4, ncol = 6)

```

##Configural invariance 

We follow the Wu and Estabrook (2016) procedure for caluclating invariance for models with variables on categorical scale
We use delta parametrization.
Groups are split based on psychiatric diagnosis (yes/no);

First we create a syntax for a configural (baseline) model using semTools package
```{r}

model3 = 'survival_strategy =~ BAPS2 + BAPS3 + BAPS4 + BAPS5 + BAPS6
        negative_beliefs =~ BAPS7 + BAPS8 + BAPS9 + BAPS10 + BAPS11 + BAPS12
        normalizing_beliefs =~ BAPS13 + BAPS14 + BAPS15 + BAPS16 + BAPS17 + BAPS18
BAPS13 ~~ BAPS14'

MI_syntax_conf = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                    ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                    group = "diagnoza", group.equal = 'configural')
```
Next, we fit the configural (baseline model).
If the fit indices are ok (Hu and Bentler, 1999) we assume configural invariance
```{r}
cfa.config = cfa(as.character(MI_syntax_conf), data = d, estimator = "WLSMV", 
                group = "diagnoza", missing = 'listwise', ordered = TRUE)
fitmeasures(cfa.config,c('chisq.scaled','df.scaled','cfi.scaled','tli.scaled',
                         'rmsea.scaled','srmr'))
```

We write the fit indices to the matrix created earlier.

```{r}
all.results[1,] =round(data.matrix((fitmeasures(cfa.config,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))), digits = 4)
```

##Treshold invariance 
First, we create a syntax for a model with thresholds constrain.
```{r}
MI_syntax_thresh = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                 group = "diagnoza", group.equal = 'thresholds')
```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.thresholds <- cfa(as.character(MI_syntax_thresh), data = d, estimator = "WLSMV", 
                      group = "diagnoza", missing = 'listwise', ordered = TRUE)
all.results[2,] =round(data.matrix((fitmeasures(cfa.thresholds,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits = 4)
```
Finally we compare baseline model and threshold model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.thresholds,cfa.config)
```
We see that the models do not differ significantly, thus we can assume threshold invariance.

##Metric invariance

First, we create a syntax for a model with thresholds and loadings constrains.

```{r}
MI_syntax_metric = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diagnoza", group.equal = c('thresholds','loadings'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.metric <- cfa(as.character(MI_syntax_metric), data = d, estimator = "WLSMV", 
                  group = "diagnoza", missing = 'listwise', ordered = TRUE)

all.results[3,] =round(data.matrix((fitmeasures(cfa.metric,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits = 4)
```
Finally we compare threshold model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.metric,cfa.thresholds)
```
Again the test was not significant so we assume metric invariance and proceed to the last step

##Scalar invariance

First, we create a syntax for a model with thresholds, loadings, and intercepts constrains.

```{r}
MI_syntax_scalar = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                                 ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diagnoza", group.equal = c('thresholds','loadings','intercepts'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.scalar <- cfa(as.character(MI_syntax_scalar), data = d, estimator = "WLSMV", 
                  group = "diagnoza", missing = 'listwise', ordered = TRUE)


all.results[4,] = round(data.matrix((fitmeasures(cfa.scalar,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits =4)
```
Finally we compare scalar model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.scalar,cfa.metric)
colnames(all.results) = c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')
print(all.results)
```
The test was not significant, thus we can assume scalar invariance. 
