---
title: "GPTS Invariance testing Joachim_Kowalski"
author: "Jan Szczypiński"
date: "08 10 2021"
output:
  pdf_document: default
  word_document: default
---

## First the MI analysis of GPTS

```{r, echo=FALSE}
library(lavaan)
library(semTools)

setwd("f:/projekty/Joachim/GPTS")
source('F:/projekty/robust_diff.R')
d <- read.csv("GPTS.csv", sep=",", header =T)
```

## recoding variables in dataset 2 and 3 
```{r setup, warning=FALSE}
library(dplyr)
d1 = subset(d, badanie ==1)
d2 = subset(d, badanie ==2)
d3 = subset(d, badanie ==3)

d2 = d2 %>% 
  mutate_at(vars(1:18), 
            ~factor(recode(.,
                           "1"=0, 
                           "2"=1,
                           "3"=2,
                           "4"=3,
                           "5"=4)),ordered=T) 

d3 = d3 %>% 
  mutate_at(vars(1:18), 
            ~factor(recode(.,
                           "1"=0, 
                           "2"=1,
                           "3"=2,
                           "4"=3,
                           "5"=4)),ordered=T) 

d = rbind(d1,d2)
d = rbind(d,d3)
colnames(d)[1] = 'PartA_gptsa1'   
```

We create a matrix to compare fit indices across models with different constraints
```{r}

all.results<-matrix(NA, nrow = 4, ncol = 6)

```

##Configural invariance 

We follow the Wu and Estabrook (2016) procedure for caluclating invariance for models with variables on categorical scale
We use delta parametrization.
Groups are split based on psychiatric diagnosis (yes/no);

First we create a syntax for a configural (baseline) model using semTools package
```{r}

model5 = 'reference =~ PartA_gptsa1 + PartA_gptsa2 +PartA_gptsa3 + PartA_gptsa4 +
          PartA_gptsa5 + PartA_gptsa6 + PartA_gptsa7 
          
          persecutory =~ partB_gptsb1 + partB_gptsb2 + partB_gptsb3 + partB_gptsb4 +
          partB_gptsb5 + partB_gptsb6 + partB_gptsb7 + partB_gptsb8 + partB_gptsb9 +
          partB_gpts10
partB_gptsb2 ~~ partB_gptsb4
partB_gptsb3 ~~ partB_gptsb4
partB_gptsb2 ~~ partB_gptsb3'

MI_syntax_conf = measEq.syntax(configural.model = model5, data = d, ordered = TRUE, 
                    ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                    group = "diag", group.equal = 'configural')
```
Next, we fit the configural (baseline model).
If the fit indices are ok (Hu and Bentler, 1999) we assume configural invariance
```{r}
cfa.config = cfa(as.character(MI_syntax_conf), data = d, estimator = "WLSMV", 
                group = "diag", missing = 'listwise', ordered = TRUE)
fitmeasures(cfa.config,c('chisq.scaled','df.scaled','cfi.scaled','tli.scaled',
                         'rmsea.scaled','srmr'))
```

We write the fit indices to the matrix created earlier.

```{r}
all.results[1,] =round(data.matrix((fitmeasures(cfa.config,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))), digits = 4)
```

##Treshold invariance 
First, we create a syntax for a model with thresholds constrain.
```{r}
MI_syntax_thresh = measEq.syntax(configural.model = model5, data = d, ordered = TRUE, 
                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                 group = "diag", group.equal = 'thresholds')
```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.thresholds <- cfa(as.character(MI_syntax_thresh), data = d, estimator = "WLSMV", 
                      group = "diag", missing = 'listwise', ordered = TRUE)
all.results[2,] =round(data.matrix((fitmeasures(cfa.thresholds,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits = 4)
```
Finally we compare baseline model and threshold model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.thresholds,cfa.config)
```
We see that the models do not differ significantly, thus we can assume threshold invariance.

##Metric invariance

First, we create a syntax for a model with thresholds and loadings constrains.

```{r}
MI_syntax_metric = measEq.syntax(configural.model = model5, data = d, ordered = TRUE, 
                                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.metric <- cfa(as.character(MI_syntax_metric), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)

all.results[3,] =round(data.matrix((fitmeasures(cfa.metric,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits = 4)
```
Finally we compare metric model and threshold model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.metric,cfa.thresholds)
```
Again the test was not significant so we assume metric invariance and proceed to the last step

##Scalar invariance

First, we create a syntax for a model with thresholds, loadings, and intercepts constrains.

```{r}
MI_syntax_scalar = measEq.syntax(configural.model = model5, data = d, ordered = TRUE, 
                                 ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings','intercepts'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.scalar <- cfa(as.character(MI_syntax_scalar), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)


all.results[4,] = round(data.matrix((fitmeasures(cfa.scalar,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits =4)
```
Finally we compare scalar model and metric model model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.scalar,cfa.metric)
colnames(all.results) = c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')
print(all.results)
```
The test was significant, thus we can't assume scalar invariance. 
The next step is to explore which constrained parameter needs to be freed to improve the fit.
We use lavTestScore function to print the parameters which have the most influence on the model
```{r}
param = lavTestScore(cfa.scalar)$uni
head(param[order(param$X2,decreasing =T),],10)
```

We need to use the parTable function to see what parameter numbers mean

```{r}
head(parTable(cfa.scalar),10)
```

We see that freeing the reference =~ PartA_gptsa2 parameter improves the model fit the most

#Partial invariance

We will free the preference =~ PartA_gptsa2 parameter and try to establish a partial scalar invariance

First, we create a syntax for a model with thresholds, loadings, and intercepts constrains.

```{r}
MI_syntax_partial = measEq.syntax(configural.model = model5, data = d, ordered = TRUE, 
                                 ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings','intercepts'), 
                                 group.partial = c('reference =~ PartA_gptsa2'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.partial <- cfa(as.character(MI_syntax_partial), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)


all.results = rbind(all.results,round((fitmeasures(cfa.partial,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))),digits =4))

```

Finally we compare partial scalar model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.partial,cfa.metric)

print(all.results)
```
We now see that there is no difference present between model with released reference =~ PartA_gptsa2 parameter
Thus, we can establish partial scalar invariance