---
title: "Joachim_CFA_report"
author: "Jan Szczypiński"
date: "08 10 2021"
output:
  pdf_document: default
  word_document: default
---

## First the analysis of GPTS

```{r, echo=FALSE}
library(lavaan)
library(semTools)

setwd("f:/projekty/Joachim/GPTS")
d <- read.csv("GPTS.csv", sep=",", header =T)
```

## recoding variables in dataset 2 and 3 
```{r setup, warning=FALSE, echo = FALSE}
library(dplyr)
d1 = subset(d, badanie ==1)
d2 = subset(d, badanie ==2)
d3 = subset(d, badanie ==3)

d2 = d2 %>% 
  mutate_at(vars(1:18), 
            ~factor(recode(.,
                           "1"=0, 
                           "2"=1,
                           "3"=2,
                           "4"=3,
                           "5"=4)),ordered=T) 

d3 = d3 %>% 
  mutate_at(vars(1:18), 
            ~factor(recode(.,
                           "1"=0, 
                           "2"=1,
                           "3"=2,
                           "4"=3,
                           "5"=4)),ordered=T) 

d = rbind(d1,d2)
d = rbind(d,d3)
colnames(d)[1] = 'PartA_gptsa1'   
```

### First model, based on previous studies (pdf): 

```{r, echo = FALSE}
library(lavaan)
model1 = 'reference =~ PartA_gptsa1 + PartA_gptsa2 +PartA_gptsa3 + PartA_gptsa4 +
          PartA_gptsa5 + PartA_gptsa6 + PartA_gptsa7  + PartA_gptsa8
          
          persecutory =~ partB_gptsb1 + partB_gptsb2 + partB_gptsb3 + partB_gptsb4 +
          partB_gptsb5 + partB_gptsb6 + partB_gptsb7 + partB_gptsb8 + partB_gptsb9 +
          partB_gpts10'

model_fit1 = cfa(model1, data = d, estimator = "WLSMV", std.lv=TRUE, missing = "listwise", ordered = T)
fitmeasures(model_fit1,c('chisq.scaled','df.scaled',                      'cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

reliability(model_fit1)

mod_indices = modificationindices(model_fit1)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)         
```

### Second model. Item 8a was removed based on fit indices from previous step of the analysis

```{r, echo = FALSE}
library(lavaan)
model2 = 'reference =~ PartA_gptsa1 + PartA_gptsa2 +PartA_gptsa3 + PartA_gptsa4 +
          PartA_gptsa5 + PartA_gptsa6 + PartA_gptsa7 
          
          persecutory =~ partB_gptsb1 + partB_gptsb2 + partB_gptsb3 + partB_gptsb4 +
          partB_gptsb5 + partB_gptsb6 + partB_gptsb7 + partB_gptsb8 + partB_gptsb9 +
          partB_gpts10'

model_fit2 = cfa(model2, data = d, estimator = "WLSMV", std.lv=TRUE, missing = "listwise", ordered = T)
fitmeasures(model_fit2,c('chisq.scaled','df.scaled',                      'cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

robust_dif = function (x,y) {
    fit_x = fitmeasures(x,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor'))
    fit_y= fitmeasures(y,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor'))
    fit_x=as.list(fit_x)
    fit_y=as.list(fit_y)
    cd = (fit_x$df.scaled * fit_x$chisq.scaling.factor - fit_y$df.scaled * fit_y$chisq.scaling.factor)/
        (fit_x$df.scaled-fit_y$df.scaled)
    Test_stat = (fit_x$chisq.scaled * fit_x$chisq.scaling.factor - 
                     fit_y$chisq.scaled * fit_y$chisq.scaling.factor) /cd
    p_val = pchisq(Test_stat,df=fit_x$df.scaled-fit_y$df.scaled, lower.tail = FALSE)
    Test_stat = round(Test_stat,digits =3)
    p_val = round(p_val,digits =3)
    Test = c('chisq.diff: ',Test_stat,'p value: ',p_val)
    return(Test)
}

reliability(model_fit2)

mod_indices = modificationindices(model_fit2)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)   
```

### Difference in robust chi-square test (see Szczypiński et al., 2021 section 2.3.2) between models 1 i 2
```{r, echo = FALSE}
library(lavaan)
robust_dif(model_fit1,model_fit2)
```

### Summary of model 2 GPTS
```{r, echo = FALSE}
summary(model_fit2, standardized=TRUE)
```


### summary of fit measures for three models of GPTSA
### deleted models 3 to 5 since they did not differ significantly from model 2
```{r, echo = FALSE}
fitmeasures(model_fit1,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
fitmeasures(model_fit2,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

```



## MUSEQ is next

### Model 1

```{r, echo = FALSE}
library(lavaan)
library(semTools)
setwd("f:/projekty/Joachim/MUSEQ")
d = read.csv("MUSEQ.csv", sep=",",  header=T)

colnames(d)[1] = 'MUSEQ_1' 


model = 'auditory =~MUSEQ_1+MUSEQ_2+MUSEQ_3+MUSEQ_4+MUSEQ_5+MUSEQ_6+MUSEQ_7

visual =~ MUSEQ2_1+MUSEQ2_2+MUSEQ2_3+MUSEQ2_4+MUSEQ2_5+MUSEQ2_6+MUSEQ2_7+MUSEQ2_8

olfactory =~ MUSEQ3_1+MUSEQ3_2+MUSEQ3_3+MUSEQ3_4+MUSEQ3_5+MUSEQ3_6+MUSEQ3_7

gustatory =~ MUSEQ4_1+MUSEQ4_2+MUSEQ4_3+MUSEQ4_4+MUSEQ4_5+MUSEQ4_6+MUSEQ4_7+MUSEQ4_8

bodily =~ MUSEQ5_1+MUSEQ5_2+MUSEQ5_3+MUSEQ5_4+MUSEQ5_5+MUSEQ5_6+MUSEQ5_7+MUSEQ5_8

presence =~ MUSEQ6_1+MUSEQ6_2+MUSEQ6_3+MUSEQ6_4

##covariances ##

'
fit = cfa(model, data = d, estimator = 'WLSMV', std.lv=TRUE, missing = "listwise", ordered = TRUE)
fitmeasures(fit,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

reliability(fit)

mod_indices = modificationindices(fit)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)
```

### Model 2 without items 3.3 i 4.3

```{r, echo = FALSE}
model2 = 'auditory =~MUSEQ_1+MUSEQ_2+MUSEQ_3+MUSEQ_4+MUSEQ_5+MUSEQ_6+MUSEQ_7

visual =~ MUSEQ2_1+MUSEQ2_2+MUSEQ2_3+MUSEQ2_4+MUSEQ2_5+MUSEQ2_6+MUSEQ2_7+MUSEQ2_8

olfactory =~ MUSEQ3_1+MUSEQ3_2+MUSEQ3_4+MUSEQ3_5+MUSEQ3_6+MUSEQ3_7

gustatory =~ MUSEQ4_1+MUSEQ4_2+MUSEQ4_4+MUSEQ4_5+MUSEQ4_6+MUSEQ4_7+MUSEQ4_8

bodily =~ MUSEQ5_1+MUSEQ5_2+MUSEQ5_3+MUSEQ5_4+MUSEQ5_5+MUSEQ5_6+MUSEQ5_7+MUSEQ5_8

presence =~ MUSEQ6_1+MUSEQ6_2+MUSEQ6_3+MUSEQ6_4

##covariances ##

'
fit2 = cfa(model2, data = d, estimator = 'WLSMV', std.lv=TRUE, missing = "listwise", ordered = TRUE)


fitmeasures(fit2,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

mod_indices = modificationindices(fit2)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)
```

### Comparison between models  1 and 2 of MUSEQ

```{r, echo = FALSE}
library(lavaan)
robust_dif(fit, fit2)
```

### Model 3 without item 6.1

```{r, echo = FALSE}
model3 = 'auditory =~MUSEQ_1+MUSEQ_2+MUSEQ_3+MUSEQ_4+MUSEQ_5+MUSEQ_6+MUSEQ_7

visual =~ MUSEQ2_1+MUSEQ2_2+MUSEQ2_3+MUSEQ2_4+MUSEQ2_5+MUSEQ2_6+MUSEQ2_7+MUSEQ2_8

olfactory =~ MUSEQ3_1+MUSEQ3_2+MUSEQ3_4+MUSEQ3_5+MUSEQ3_6+MUSEQ3_7

gustatory =~ MUSEQ4_1+MUSEQ4_2+MUSEQ4_4+MUSEQ4_5+MUSEQ4_6+MUSEQ4_7+MUSEQ4_8

bodily =~ MUSEQ5_1+MUSEQ5_2+MUSEQ5_3+MUSEQ5_4+MUSEQ5_5+MUSEQ5_6+MUSEQ5_7+MUSEQ5_8

presence =~ MUSEQ6_2+MUSEQ6_3+MUSEQ6_4

##covariances ##

'
fit3 = cfa(model3, data = d, estimator = 'WLSMV', std.lv=TRUE, missing = "listwise", ordered = TRUE)

fitmeasures(fit3,c('chisq.scaled','df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

reliability(fit3)


```

### Comparison between models 2 and 3 of MUSEQ

```{r, echo = FALSE}
library(lavaan)
robust_dif(fit2, fit3)
```
### Summary of model 3 MUSEQ
```{r, echo = FALSE}
summary(fit3, standardized=TRUE)
```
### summary of fit measures for three models of MUSEQ

```{r, echo = FALSE}
fitmeasures(fit,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
fitmeasures(fit2,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
fitmeasures(fit3,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
```

## Analiza BAPS

```{r, echo = FALSE}
setwd("f:/projekty/Joachim/BAPS")

d <- read.csv("BAPS.csv", sep=",", header =T)

library(lavaan)
library(semTools)
colnames(d)[1] = 'BAPS1' 
```
### 1st model
```{r, echo = FALSE}
model = 'survival_strategy =~ BAPS1 + BAPS2 + BAPS3 + BAPS4 + BAPS5 + BAPS6
        negative_beliefs =~ BAPS7 + BAPS8 + BAPS9 + BAPS10 + BAPS11 + BAPS12
        normalizing_beliefs =~ BAPS13 + BAPS14 + BAPS15 + BAPS16 + BAPS17 + BAPS18'

fit = cfa(model, data = d, estimator = 'WLSMV', ordered = TRUE, std.lv=TRUE, missing = "listwise",)

fitmeasures(fit,c('chisq.scaled','df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

reliability(fit)

mod_indices = modificationindices(fit)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)

```

### Model 2 -  without BAPS1
```{r, echo = FALSE}
model2 = 'survival_strategy =~ BAPS2 + BAPS3 + BAPS4 + BAPS5 + BAPS6
        negative_beliefs =~ BAPS7 + BAPS8 + BAPS9 + BAPS10 + BAPS11 + BAPS12
        normalizing_beliefs =~ BAPS13 + BAPS14 + BAPS15 + BAPS16 + BAPS17 + BAPS18'

fit2 = cfa(model2, data = d, estimator = 'WLSMV', ordered = TRUE, std.lv=TRUE, missing = "listwise",)

fitmeasures(fit2,c('chisq.scaled','df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))

reliability(fit2)

mod_indices = modificationindices(fit2)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)

```

### Comparison between model 1 and 2 of BAPS

```{r, echo = FALSE}
library(lavaan)
robust_dif(fit, fit2)
```

#### MODEL 3 with covariance added between BAPS13 i BAPS14
```{r, echo = FALSE}
model3 = 'survival_strategy =~ BAPS2 + BAPS3 + BAPS4 + BAPS5 + BAPS6
        negative_beliefs =~ BAPS7 + BAPS8 + BAPS9 + BAPS10 + BAPS11 + BAPS12
        normalizing_beliefs =~ BAPS13 + BAPS14 + BAPS15 + BAPS16 + BAPS17 + BAPS18
BAPS13 ~~ BAPS14'

fit3 = cfa(model3, data = d, estimator = 'WLSMV', ordered = TRUE, std.lv=TRUE, missing = "listwise")
fitmeasures(fit3,c('chisq.scaled','df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
reliability(fit3)
```

### Comparison between model 2 and 3 of BAPS

```{r, echo = FALSE}
library(lavaan)
robust_dif(fit2, fit3)
```

### Summary of model 3 BAPS
```{r, echo = FALSE}
summary(fit3, standardized=TRUE)
```

### summary of fit measures for three models of BAPS

```{r, echo = FALSE}
fitmeasures(fit,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
fitmeasures(fit2,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
fitmeasures(fit3,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
```

## IVI analysis

```{r, echo = FALSE}
setwd("f:/projekty/Joachim/IVI")

d <- read.csv("IVI.csv", sep=",", header =T)

library(lavaan)
library(semTools)
colnames(d)[1] = 'IVI_1' 

model = 'metaphysical =~ IVI_1 + IVI_3 + IVI_4 + IVI_5 + IVI_6 + IVI_8 + IVI_13 +
IVI_14 + IVI_15 + IVI_16 + IVI_17 + IVI_25 + IVI_26
positive =~ IVI_2 + IVI_7 + IVI_9 +IVI_18 +IVI_19 +IVI_20 + IVI_21 + IVI_24
loss_of_control =~ IVI_10 +IVI_11 +IVI_12 +IVI_22 +IVI_23'

fit = cfa(model, data = d, estimator = 'WLSMV', ordered = TRUE, std.lv=TRUE, missing = "listwise")
summary(fit,standardized=T, fit.measures = T)

```


### the model was not identified, so the the factor solution was not optimal.  Thus we did run an iclust algorithm to find alternative factor model
### the iclust plot is kind of hard to read so I saved it in separate .pdf file "iclust_ivi.pdf"
``` {r, echo = FALSE}
library(psych)
library(EFA.dimensions)
iclust((POLYCHORIC_R(d, method = "Fox", verbose = F)))
```

## PQ16 analysis, starting with simple 1-factor model, only with data from 2nd study
```{r, echo = FALSE}
setwd("f:/projekty/Joachim/PQ16")

d <- read.csv("PQ16.csv", sep=",", header =T)

library(lavaan)
library(semTools)

colnames(d)[1] = 'pq1' 

d = subset(d, badanie == 2)

model1 = ' factor =~ pq1 + pq2 + pq3 + pq4 + pq5 + pq6 + pq7 + pq8 + pq9 + pq10 +
                    pq11 + pq12 + pq13 + pq14 + pq15 + pq16'
fit1 = cfa(model1, data = d, estimator = 'WLSMV', std.lv=TRUE, missing = "listwise", ordered = TRUE)


fitmeasures(fit1,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))


reliability(fit1)

mod_indices = modificationindices(fit1)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)
```

### Next, we added covariance between pq7 and pq14

```{r, echo = FALSE}

model2 = ' factor =~ pq1 + pq2 + pq3 + pq4 + pq5 + pq6 + pq7 + pq8 + pq9 + pq10 +
                    pq11 + pq12 + pq13 + pq14 + pq15 + pq16
pq7 ~~ pq14'
fit2 = cfa(model2, data = d, estimator = 'WLSMV', std.lv=TRUE, missing = "listwise", ordered = TRUE)


fitmeasures(fit2,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))


reliability(fit2)

mod_indices = modificationindices(fit2)
head(mod_indices[order(mod_indices$mi, decreasing = T),],10)
```

### Comparison between model 1 and 2 of PQ16

```{r, echo = FALSE}
library(lavaan)
robust_dif(fit1, fit2)
```


### and in the next model, we added covariance between pq1 and pq7

```{r, echo = FALSE}
model3 = ' factor =~ pq1 + pq2 + pq3 + pq4 + pq5 + pq6 + pq7 + pq8 + pq9 + pq10 +
                    pq11 + pq12 + pq13 + pq14 + pq15 + pq16
pq7 ~~ pq14
pq1 ~~ pq7'
fit3 = cfa(model3, data = d, estimator = 'WLSMV', std.lv=TRUE, missing = "listwise", ordered = TRUE)


fitmeasures(fit3,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))


reliability(fit3)
```

### Comparison between model 1 and 2 of PQ16

```{r, echo = FALSE}
library(lavaan)
robust_dif(fit2, fit3)
```

### Summary of model 3 PQ16
```{r, echo = FALSE}
summary(fit3, standardized=TRUE)
```

### summary of fit measures for three models of PQ16

```{r, echo = FALSE}
fitmeasures(fit1,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
fitmeasures(fit2,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
fitmeasures(fit3,c('chisq.scaled', 'df.scaled', 'chisq.scaling.factor','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))
```