---
title: "PQ16 Invariance testing Joachim_Kowalski"
author: "Jan Szczypiński"
date: "08 10 2021"
output:
  pdf_document: default
  word_document: default
---

## The MI analysis of PQ16

First loading the packages, data and a function.
We use data only from study 2
```{r, echo=FALSE}
setwd("f:/projekty/Joachim/PQ16")
d <- read.csv("PQ16.csv", sep=",", header =T)
source('F:/projekty/robust_diff.R')
library(lavaan)
library(semTools)

colnames(d)[1] = 'pq1' 

d = subset(d, badanie == 2)
```

We create a matrix to compare fit indices across models with different constraints
```{r}

all.results<-matrix(NA, nrow = 4, ncol = 6)

```

##Configural invariance 

We follow the Wu and Estabrook (2016) procedure for caluclating invariance for models with variables on categorical scale
We use delta parametrization.
Groups are split based on psychiatric diagnosis (yes/no);

First we create a syntax for a configural (baseline) model using semTools package
```{r}

model3 = ' factor =~ pq1 + pq2 + pq3 + pq4 + pq5 + pq6 + pq7 + pq8 + pq9 + pq10 +
                    pq11 + pq12 + pq13 + pq14 + pq15 + pq16
pq7 ~~ pq14
pq1 ~~ pq7'

MI_syntax_conf = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                    ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                    group = "diag", group.equal = 'configural')
```
Next, we fit the configural (baseline model).
If the fit indices are ok (Hu and Bentler, 1999) we assume configural invariance
```{r}
cfa.config = cfa(as.character(MI_syntax_conf), data = d, estimator = "WLSMV", 
                group = "diag", missing = 'listwise', ordered = TRUE)
fitmeasures(cfa.config,c('chisq.scaled','df.scaled','cfi.scaled','tli.scaled',
                         'rmsea.scaled','srmr'))
```

We write the fit indices to the matrix created earlier.

```{r}
all.results[1,] =round(data.matrix((fitmeasures(cfa.config,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))), digits = 4)
```

##Treshold invariance 
First, we create a syntax for a model with thresholds constrain.
```{r}
MI_syntax_thresh = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                 group = "diag", group.equal = 'thresholds')
```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.thresholds <- cfa(as.character(MI_syntax_thresh), data = d, estimator = "WLSMV", 
                      group = "diag", missing = 'listwise', ordered = TRUE)
all.results[2,] =round(data.matrix((fitmeasures(cfa.thresholds,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits = 4)
```
Finally we compare baseline model and threshold model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.thresholds,cfa.config)
```
We see that the models do not differ significantly, thus we can assume threshold invariance.

##Metric invariance

First, we create a syntax for a model with thresholds and loadings constrains.

```{r}
MI_syntax_metric = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.metric <- cfa(as.character(MI_syntax_metric), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)

all.results[3,] =round(data.matrix((fitmeasures(cfa.metric,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits = 4)
```
Finally we compare threshold model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.metric,cfa.thresholds)
```
The test was significant so we can't assume metric invariance
The next step is to explore which constrained parameter needs to be freed to improve the fit.
We use lavTestScore function to print the parameters which have the most influence on the model
```{r}
param = lavTestScore(cfa.metric)$uni
head(param[order(param$X2,decreasing =T),],10)
```

We need to use the parTable function to see what parameter numbers mean

```{r}
head(parTable(cfa.metric),10)
```
We will free the factor	=~	pq1 parameter to see if we can get partial metric invariance
```{r}
MI_syntax_metric.part = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings'), group.partial = 'factor	=~	pq1')

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.metric.part <- cfa(as.character(MI_syntax_metric.part), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)

all.results[4,] =round(data.matrix((fitmeasures(cfa.metric,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr')))),digits = 4)
```
Finally we compare threshold model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.metric.part,cfa.thresholds)
```
The test is still significant so we will look for another parameter
```{r}
param = lavTestScore(cfa.metric.part)$uni
head(param[order(param$X2,decreasing =T),],10)
```

We need to use the parTable function to see what parameter numbers mean

```{r}
head(parTable(cfa.metric.part),10)
```
We will free the factor	=~	pq11 parameter to see if we can get partial metric invariance
```{r}
MI_syntax_metric.part = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings'), 
                                group.partial = c('factor	=~	pq1','factor	=~	pq11'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.metric.part <- cfa(as.character(MI_syntax_metric.part), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)

all.results =rbind(all.results,round((fitmeasures(cfa.metric,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))),digits = 4))
```
Finally we compare threshold model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.metric.part,cfa.thresholds)

```
We will free the factor	=~	pq7 parameter to see if we can get partial metric invariance
```{r}
MI_syntax_metric.part = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings'), 
                                group.partial = c('factor	=~	pq1','factor	=~	pq11','factor	=~	pq7'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.metric.part <- cfa(as.character(MI_syntax_metric.part), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)

all.results =rbind(all.results,round((fitmeasures(cfa.metric,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))),digits = 4))
```
Finally we compare threshold model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.metric.part,cfa.thresholds)
```

We will free the factor	=~	pq2 parameter to see if we can get partial metric invariance
```{r}
MI_syntax_metric.part = measEq.syntax(configural.model = model3, data = d, ordered = TRUE, 
                                ID.cat ='Wu.2016', ID.fac ='unit.variance', parameterization = 'delta',
                                 group = "diag", group.equal = c('thresholds','loadings'), 
                                group.partial = c('factor	=~	pq1','factor	=~	pq11','factor	=~	pq7','factor=~pq2'))

```
Next, we fit the model and  save  fit indices to the matrix

```{r}
cfa.metric.part <- cfa(as.character(MI_syntax_metric.part), data = d, estimator = "WLSMV", 
                  group = "diag", missing = 'listwise', ordered = TRUE)

all.results =rbind(all.results,round((fitmeasures(cfa.metric,c('chisq.scaled',
                'df.scaled','cfi.scaled','tli.scaled','rmsea.scaled','srmr'))),digits = 4))
```
Finally we compare threshold model and metric model using Satorra-Bentler Scaled chi^2 difference test
```{r}
robust_diff(cfa.metric.part,cfa.thresholds)
```
Even with 4 freely estimated factor loadings the partial metric invariance still couldn't be met. Thus, we conclude that metric invariance is not present for the pq16
